### 2017-01-31 doc/manuscript.Rmd
Added Bernie onto author list
Moved degree before affil number
Moved more detailed GEE stats methods to Supplemental Material

Author: Luke W. Johnston (1), MSc; Stewart B. Harris (2), MD; Ravi Retnakaran
(3,4), MD; Adria Giacca (5), PhD; Zhen Liu (1), PhD;
Richard P. Bazinet (1), PhD; and Anthony J. Hanley (1,6), PhD

@@

Generalized estimating equation (GEE) models [@Zeger1986a] were used in the
primary analysis to determine the longitudinal associations between the outcome
variables and the predictor variables. Given the longitudinal design, an
auto-regressive of order 1 (AR1) working correlation matrix was chosen for the GEE
models as it had the best model fit when assessed using quasi-likelihood
information criteria (QIC), though other matrices (eg. exchangeable)
had similar fit (data not shown). GEE is well suited to longitudinal cohort 
studies given it's capacity to handle missed visits. The predictor variables and
continuous covariates were scaled (mean-centered and standardized). The TAGFA,
total NEFA, sex, ethnicity, and baseline age were classified as 
*time-independent* (held constant) as they were measured only at the baseline 
visit or do not change throughout the study, while the outcome variables and 
remaining covariates were set as *time-dependent*. No imputation was conducted 
on missing values. Given how the samples for TAGFA measurement were determined
and how the outcome variables are calculated, cases of diabetes at baseline and
followup were not included in the GEE modeling.

Covariates were selected based on the previous literature, from directed acyclic
graph [@Greenland1999a] recommendations (using the DAGitty software
[@Textor2011a], http://dagitty.net/; see `r cite_f('dag_is')` and `r cite_f('dag_bcf')`), and from
QIC. `r cite_t('qic')` shows the covariates compared using QIC.
While the final GEE model selected as best fitting differed between insulin
sensitivity and beta-cell function measures, the differences in QIC values were
less than 10 between many of the models, suggesting similar fit. As such, we selected
the model that had the fewest covariates and that had similar fit between the
outcome measures. The final GEE model was adjusted for time, waist
circumference, baseline age, ethnicity, sex, ALT, MET, and total NEFA.
After scaling, log-transforming, and exponentiating, the GEE estimates are 
interpreted as an expected percent difference in the outcome variable for every 
standard deviation (SD) increase in the predictor variable given the covariates 
are held constant (including time). Lastly for the GEE models, we tested for an
interaction with sex, ethnicity, or time by the predictor term for each outcome
variable.

While GEE accounts for the longitudinal design of the data, this approach is 
limited in that it cannot analyze the inherent multivariate nature of the 
composition of the TAGFA fraction. To confirm the GEE results in a multivariate 
environment (all TAGFA), partial least squares regression (PLS) was
used to identify the patterns of TAGFA composition against insulin sensitivity
and beta-cell function. No prevalent diabetes cases were included in the PLS
analysis. For more detailed explanation of GEE and PLS, please see the
{{supplemental}} Methods.

@@

>>>>

Author: Luke W. Johnston, MSc (1); Stewart B. Harris, MD (2); Ravi Retnakaran, MD
(3,4); Bernard Zinman, MD (3,4); Adria Giacca, PhD (5); Zhen Liu, PhD (1);
Richard P. Bazinet, PhD (1); and Anthony J. Hanley, PhD (1,6)

@@

Generalized estimating equation (GEE) models [@Zeger1986a] were used in the 
primary analysis to determine the longitudinal associations between the outcome 
variables and the predictor variables. The predictor variables and continuous 
covariates were scaled (mean-centered and standardized). Given the longitudinal 
design, an auto-regressive of order 1 (AR1) working correlation matrix was 
specified in the GEE model. Covariates to adjust for were selected based on the
previous literature, from directed acyclic graph [@Greenland1999a]
recommendations, and from quasi-likelihood information criteria (QIC). The final
GEE model was adjusted for time, waist circumference, baseline age, ethnicity,
sex, ALT, MET, and total NEFA. The TAGFA, total NEFA, sex, ethnicity, and
baseline age were classified as *time-independent* (held constant) as they were
measured only at the baseline visit or do not change throughout the study, while
the outcome variables and remaining covariates were set as *time-dependent*.
After transformations, the GEE estimates are interpreted as an expected percent 
difference in the outcome variable for every standard deviation (SD) increase in
the predictor variable given the covariates are held constant (including time). 
We also tested for an interaction with sex, ethnicity, or time by the predictor 
term for each outcome variable. See the Supplemental Methods for an expanded 
explanation of the GEE modeling analysis.

While GEE accounts for the longitudinal design of the data, this approach is 
limited in that it cannot analyze the inherent multivariate nature of the 
composition of the TAGFA fraction. To confirm the GEE results in a multivariate 
environment (all TAGFA), partial least squares regression (PLS) was
used to identify the patterns of TAGFA composition against insulin sensitivity
and beta-cell function. No prevalent diabetes cases were included in the PLS
analysis. For more detailed explanation of PLS, please see the Supplemental
Methods.

@@

## Supplemental Methods: Statistical analysis

Generalized estimating equations (GEE) is a technique similar to mixed effects
modeling, except it calculates the marginal population averages compared to the
individual averages in mixed effects models. GEE is well suited and commonly
used in longitudinal cohort studies, especially given it's capacity to handle
missed visits.

The working correlation matrix for the GEE analysis was chosen based on 
quasi-likelihood information criteria (QIC). The auto-regressive of order 1
(AR1) matrix was chosen for the GEE models as it had the best model fit accessed
using QIC, though other matrices (eg. exchangeable) had similar fit (data not
shown).

For the confounders, they were chosen based on previous literature, from
directed acyclic graph (DAG) [@Greenland1999a] recommendations, and from
QIC. The DAG recommendations were obtained from using the DAGitty software
[@Textor2011a], http://dagitty.net/. See `r cite_f('dag_is')` and 
`r cite_f('dag_bcf')`) for the DAG model. See `r cite_t('qic')` for the
comparison of various models with different covariates using QIC. While the
final GEE model selected as best fitting differed between insulin sensitivity
and beta-cell function measures, the differences in QIC values were less than 10
between many of the models, suggesting similar fit. As such, we selected the
model that had the fewest covariates and that had similar fit between the
outcome measures.

No imputation was conducted on missing values. Given how the samples for TAGFA
measurement were determined and how the outcome variables are calculated, cases
of diabetes at baseline and followup were not included in the GEE modeling. The
resulting GEE beta estimates were exponentiated to allow the interpretation as
stated in the main methods.
